ALIGNMENT OF LIGHT AND MASS IN LENSING GALAXIES

### 2013-02-20 shapes/thesis.tex
thesis update

\subsection{Anything done before?}
\textbf{\cite{1997ApJ...482..604K}}: Look at 17 lenses at redshifts ranging between 0.1 and $\sim1$. Sample consists of mainly passively evolving elliptical galaxies and a few spirals. Claim that generally projected light and dark matter distributions are strongly correlated and are not far out of alignment ($\lesssim10^{\circ}$), except in cases where there are strong external tidal perturbations (although misalignments are not necessary in these cases). 'Dark matter halos can be significantly flatter than the light (e.g., Sackett et al. 1994; Buote \& Canizares 1994, 1996; also see the review by Sackett 1996)'. To reconstruct the mass they fit singular isothermal ellipsoid (SIE) lens models, which immediately give them the shape. Another model they use is a singular isothermal sphere (SIS) and some external tidal perturbation (shear). As already stated in \cite{1998ApJ...509..561K}, these models sometimes give a bad fit. To obtain a good fit, an another independent source of external shear is needed (external tidal perturbation). They also state that it is known that the lenses Q0957 + 561 (Young et al. 1980, Grogin \& Narayan 1996), PG 1115 + 080 (Keeton \& Kochanek 1997, Kundic et al. 1997a, Schechter et al. 1997, Tonry 1998), B1422 + 231 (Hogg \& Blandford 1994, Kormann et al. 1994, Kundic et al 1997b, Tonry 1998) are subject to significant tidal perturbations by a nearby cluster, group or galaxy (also line-of-sight structure) (--> All lying in a environment? Ask Prasenjit).

@@

\textbf{\cite{2000ApJ...538L.113N}}:

\textbf{\cite{2002sgdh.conf...62K}}:

\textbf{\cite{2005MNRAS.360.1333W}}:

\textbf{\cite{2008MNRAS.383..857F}}:

@@

\textbf{\cite{2012ApJ...761..170G}}:

@@

\section{Data}
\begin{itemize}
\item SPS (Ignacio Ferreras, Dominik, etc. (\cite{2011ApJ...740...97L}))
\item Bundle information in tables (especially what input data was used for the lenses)
\end{itemize}

>>>>

\subsection{Anything done before?}
\textbf{\cite{1997ApJ...482..604K}}: Look at 17 lenses at redshifts ranging between 0.1 and $\sim1$. Sample consists of mainly passively evolving elliptical galaxies and a few spirals. Claim that generally projected light and dark matter distributions are strongly correlated and are not far out of alignment ($\lesssim10^{\circ}$), except in cases where there are strong external tidal perturbations (although misalignments are not necessary in these cases). 'Dark matter halos can be significantly flatter than the light (e.g., Sackett et al. 1994; Buote \& Canizares 1994, 1996; also see the review by Sackett 1996)'. To reconstruct the mass they fit singular isothermal ellipsoid (SIE) lens models, which immediately give them the shape. Another model they use is a singular isothermal sphere (SIS) and some external tidal perturbation (shear). As already stated in \cite{1998ApJ...509..561K}, these models sometimes give a bad fit. To obtain a good fit, an another independent source of external shear is needed (external tidal perturbation). They also state that it is known that the lenses Q0957 + 561 (Young et al. 1980, Grogin \& Narayan 1996), PG 1115 + 080 (Keeton \& Kochanek 1997, Kundic et al. 1997a, Schechter et al. 1997, Tonry 1998), B1422 + 231 (Hogg \& Blandford 1994, Kormann et al. 1994, Kundic et al 1997b, Tonry 1998) are subject to significant tidal perturbations by a nearby cluster, group or galaxy (also line-of-sight structure) (--> All lying in a environment? Ask Prasenjit). They find no obvious correlation in the axis ratios of the distributions.

@@

@@

\textbf{\cite{2006ApJ...649..599K}}: They use 15 massive field early-type galaxies selected from the SLACS Survey. To reconstruct the mass distribution (for total mass) they use a SIE mass model (Kormann et al. 1994) without external shear (they justify it by stating that an external shear was not required and also in hindsight as a small $\Delta\theta$ rms spread was reconstructed). The light distribution is based on the nonparametric source reconstruction method by Warren & Dye 2003 and is described in Treu & Koopmans 2004 and Koopmans 2005. They analyze then the misalignment between light and mass and find that the average difference is $\left<\Delta\theta\right>=0^{\circ}\pm3^{\circ}$ with a rms spread of $10^{\circ}$. For almost spherical SIEs, i.e. $q_{\mathrm{SIE}}>0.75$ the rms spread is $13^{\circ}$, for less spherical distributions, $q_{\mathrm{SIE}}<0.75$, it is only $3^{\circ}$. They conclude that this implies that dark matter is aligned with the stellar component on scales $\lesssim4$ kpc. They also find using a theoretical formula for the external shear from \cite{1997ApJ...482..604K} and assuming that there is no correlation between the galaxies and external shear orientations, they find that the average external shear has an upper limit of $\left<\gamma_{\mathrm{ext}}\right>\lesssim0.035$. They conclude therefore that the alignment of mass and light confirms that the external shear is very small and can be neglected. They state that significant external shear would in general cause a spread in $\Delta\theta$ if not accounted for in the models.

\textbf{\cite{2009ApJ...690..670T}}: They look at 70 lenses from the SLACS Survey, mostly early-type galaxies (62), 6 disk galaxies, and 2 have ambiguos morphology. They again fit a SIE without external shear and compare the position angles of the distributions. They argue that the stellar contribution is significant at the scales they test and therefore they expect no misalignment if there is no significant external potential. Also the contrary should hold that the rms amplitude of the misalignment is a measure of the external perturbing potential (\cite{1997ApJ...482..604K}; \cite{2006ApJ...649..599K}). They find that the correlation between alignment and local resp. global environment is significant. They do not expect a simple relation between these. However, in general lenses in overdense environments will be more likely to be affected by nearby companions, which causes a misalignment (external perturber located in a random direction). They estimate again that external shear of order 0.05-0.06 is needed to match the results.

\textbf{\cite{2012ApJ...761..170G}}: They use 16 massive early-type galaxies from SL2S (only 2 are quads). They reconstruct the lens mass with a SIE. They only add an external shear when more information is available. To find the intrinsic source light distribution they try to minimize a $\chi^{2}$ (they have an observed, pixelated light distribution). All in all, they use a MCMC technique to find the best-fitting 9-parameter-model  (6 from light, 3 from SIE (ellipticity, position angle, Einstein radius). They find that stellar and mass ellipticities are tightly correlated (think it is reasonable as light makes up a significant part) with the exception of a few outliers (some basically spherical mass distributions, and one case where the stars are almost round and the mass is flatte -> state that source of ellipticities seems to be external shear associated with a satellite). The scatter they find in the ellipticities ratio is however larger than it is in the SLACS sample, which is also at a lower redshift. This is probably due to a much more significant role played by external shear. They also find reasonable alignment between light and mass, the scatter is however larger than for the SLACS sample. For almost round mass distributions the scatter is $25^{\circ}$, for less round distributions it is still $18^{\circ}$. Again the calculate the required average external shear and find (independent of including the round lenses or not) $\left<\gamma_{\mathrm{ext}}\right>=0.12\pm0.05$. This is significantly larger than what they found for the SLACS lenses. They think that it is likely due to the environment of the lenses (\cite{2007AJ....134..668A}; \cite{2009ApJ...690..670T}; Auger 2008; \cite{2011ApJ...726...84W}).

@@

\section{Data}
\begin{itemize}
\item SPS (Ignacio Ferreras, Dominik, etc. (\cite{2011ApJ...740...97L}))
\item Write why specifically these lenses were chosen, where they're from
\item Bundle information in tables (especially what input data was used for the lenses)
\end{itemize}
